---
title: "BasicProphagePlots"
output: html_document
---

```{r}
library(ggrepel)

```


```{r}
#NOTE: Run DataWrangling before this


#Creates a numeric ID value instead of a character, for re-ordering
prophage_plot <- data_bacterial_viral%>%
  mutate(isolateID_num = as.numeric(isolateID))

#Replaces NA values in quality check to indicate no virus
prophage_plot$checkv_checkv_quality <- prophage_plot$checkv_checkv_quality%>% replace_na("No Virus")

#Counts the number of viruses in an isolate
virus_count <- prophage_plot%>%
  group_by(isolateID)%>%
  dplyr::count(is_virus)%>%
  filter(is_virus=="yes")%>%
  select(isolateID, n)


#Attaches virus number to initial dataframe
prophage_plot2 <- prophage_plot %>%
left_join(virus_count, by="isolateID")%>%
  mutate(n = coalesce(n, 0))%>%
  mutate(colortext = case_when( #new variable for coloring
           n > 0 ~ "some",
           n == 0 ~ "nada"
         ))

```


Table with basic viral info:
```{r}
basic_viral_table<-prophage_plot2 %>%
  filter(is_virus=="yes")%>%
  select(isolateID, bio_name, bp_length, checkv_completeness)

write.csv(basic_viral_table,"basic_viral_table.csv", row.names = TRUE)

```


Plots with isolate ID v checkv completeness/contamination: nedium, high and complete checkv quality
```{r}
#Plot with isolate ID v checkv completeness: medium, high and complete
ggplot(prophage_plot2,
       aes(x = fct_reorder(isolateID, isolateID_num),
                          y = checkv_completeness))+
  geom_point(alpha=0.5,
             size=4,
             color="blue")+
  coord_cartesian(ylim = c(0, 105))+
  labs(x = "Isolate ID",
       y = "checkv Phage Completeness (%)")+
  geom_text(aes(label= n,
                color=colortext),
          y=105,
          stat='count',
          size=4)+
  scale_color_manual(values = c('some' = 'chartreuse4', 'nada' = 'red'),
                     guide = "none")+
  ggtitle("Phage completeness per isolate (medium quality, high quality, complete)")

#Note: Warning is ok, just referring to NA values in completeness when there are no viruses

#Plot with isolate ID v checkv contamination: medium, high and complete
ggplot(prophage_plot2,
       aes(x = fct_reorder(isolateID, isolateID_num),
                          y = checkv_contamination))+
  geom_point(alpha=0.5,
             size=4,
             color="blue")+
  coord_cartesian(ylim = c(0, 105))+
  labs(x = "Isolate ID",
       y = "checkv Phage Contamination (%)")+
  geom_text(aes(label= n,
                color=colortext),
          y=105,
          stat='count',
          size=4)+
  scale_color_manual(values = c('some' = 'chartreuse4', 'nada' = 'red'),
                     guide = "none")+
  ggtitle("Phage contamination per isolate (medium quality, high quality, complete)")

#Note: Warning is ok, just referring to NA values in contamination when there are no viruses

```

Plot with isolate ID v checkv completeness, only high and complete checkv quality:
```{r}
#Filters by high quality or complete viral genomes
only_high_quality <-prophage_plot%>%
  filter(is.na(checkv_checkv_quality)| !str_detect(checkv_checkv_quality, 'Medium-quality'))%>%
  group_by(isolateID)

only_high_quality


#Second viral count with only high quality viruses
virus_count2 <- only_high_quality%>%
    dplyr::count(is_virus)%>%
  filter(is_virus=="yes")%>%
  select(isolateID, n)


#Attaches virus number to initial dataframe
only_high_quality <- only_high_quality %>%
  left_join(virus_count2, by="isolateID")%>%
  mutate(n = coalesce(n, 0))%>%
  mutate(colortext = case_when( #new variable for coloring
           n > 0 ~ "some",
           n == 0 ~ "nada"
         ))


#Plot with isolate ID v checkv completeness: high and complete checkv quality only
ggplot(only_high_quality,
       aes(x = fct_reorder(isolateID, isolateID_num),
                          y = checkv_completeness))+
  geom_point(alpha=0.5,
             size=4,
             color="blue")+
  ylim(min=0,
       max=105) +
  labs(x = "Isolate ID",
       y = "checkv Phage Completeness (%)")+
  geom_text(aes(label= n,
                color=colortext),
          y=105,
          stat='count',
          size=4)+
  scale_color_manual(values = c('some' = 'chartreuse4', 'nada' = 'red'),
                     guide = "none")+
  ggtitle("Phage completeness per isolate (high quality and complete checkv quality only)")

#Note: Warning is ok, just referring to NA values in contamination when there are no viruses



#Plot with isolate ID v checkv contamination: high and complete checkv quality only
ggplot(only_high_quality,
       aes(x = fct_reorder(isolateID, isolateID_num),
                          y = checkv_contamination))+
  geom_point(alpha=0.5,
             size=4,
             color="blue")+
  ylim(min=0,
       max=105) +
  labs(x = "Isolate ID",
       y = "checkv Phage Contamination (%)")+
  geom_text(aes(label= n,
                color=colortext),
          y=105,
          stat='count',
          size=4)+
  scale_color_manual(values = c('some' = 'chartreuse4', 'nada' = 'red'),
                     guide = "none")+
  ggtitle("Phage contamination per isolate (high quality and complete checkv quality only)")

#Note: Warning is ok, just referring to NA values in contamination when there are no viruses


```


```{r}

#Leaves only one isolate row, if they have multiple phages 
water_plot <- prophage_plot2 %>%
  distinct(isolateID, .keep_all = TRUE)

#Plot
ggplot(water_plot,
       aes(x = water_origin, y = salinity, color=is_virus))+
  geom_tile(aes(fill = water_origin,
                alpha = 0.99),
            color = "white")+
    geom_point(position = position_jitter(w=0.2, 
                                        h =0, 
                                        seed=1),
             color="black")+
  geom_label_repel(aes(label=isolateID), 
                  position = position_jitter(w=0.2, h=0, seed=1),
                  min.segment.length = unit(0, 'lines'),
                  max.overlaps = 20)+
  labs(title = "Title")+
  xlab ("Original Source")+
  ylab ("Media Salinity (g/L)")+
  scale_fill_manual(values=c("cyan3",
                             "blue"),
                    guide="none")+
  scale_color_manual(values=c("brown3",
                              "darkgreen"),
                     name= "Virus",
                     label = c("No", "Yes"))+
  guides(alpha = "none",
         water_origin="none")+
  ggtitle("Original Water Source v Media Salinity")

 
```
Bacteriophage per genus plots:
```{r}
#Quality per genus of the phage genomes
ggplot(prophage_plot2, aes(x = checkv_checkv_quality, fill = genus))+
  geom_bar(position="stack")


#Finds out which isolates have at least one virus
genus_check <- prophage_plot2%>%
  distinct(isolateID, .keep_all = TRUE)%>%
  filter(n > 0)

#Plot of genera with at least one virus
ggplot(genus_check, aes(x = genus,
                        fill=genus))+
  geom_bar()+
  theme(axis.text.x = element_blank())+
  xlab("Genus")+
  ylab("Number of isolates with at least one phage")


#Plot of family with at least one virus
ggplot(genus_check, aes(x = family,
                        fill=genus))+
  geom_bar()+
  xlab("Family")+
  ylab("Number of isolates with at least one phage")+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))


prophage_plot2%>%
  filter(Completeness < 70)%>%
  filter(is_virus=="yes")

```


Comparing STENOTROPHOMONAS sp.:
```{r}
steno <- prophage_plot2%>%
  filter(genus =="Stenotrophomonas")%>%
  filter(is_virus =="yes")

ggplot(steno,
       aes(x=isolateID,
           y=bp_length,
           color=bio_name,
           shape=water_origin))+
  geom_point(aes(alpha=0.5,
                 size=1))+
  xlab("Stenotrophomonas Isolate ID")+
  ylab("Prophage Length (bp)")+
  labs(color = "Species")+
  guides(alpha = "none", size="none")

```

Comparing VIBRIO sp.:
```{r}
vibrio <- prophage_plot2%>%
  filter(genus =="Vibrio")%>%
  filter(is_virus =="yes")

ggplot(vibrio,
       aes(x=isolateID,
           y=bp_length,
           color=bio_name,
           shape=water_origin))+
  geom_point(aes(alpha=0.5,
                 size=1))+
  xlab("Vibrio sp. Isolate ID")+
  ylab("Prophage Length (bp)")+
  labs(color = "Species")+
  guides(alpha = "none", size="none", shape="none")

```

Comparing ENTEROBACTER sp.:
```{r}
entero <- prophage_plot2%>%
  filter(genus =="Enterobacter")%>%
  filter(is_virus =="yes")

ggplot(entero,
       aes(x=isolateID,
           y=bp_length,
           color=bio_name,
           shape=water_origin))+
  geom_point(aes(alpha=0.5,
                 size=1))+
  xlab("Enterobacter sp. Isolate ID")+
  ylab("Prophage Length (bp)")+
  labs(color = "Species")+
  guides(alpha = "none", size="none", shape="none")

```


Comparing PROVIDENCIA sp.:
```{r}
providencia <- prophage_plot2%>%
  filter(genus =="Providencia")%>%
  filter(is_virus =="yes")

providencia

ggplot(providencia,
       aes(x=isolateID,
           y=bp_length,
           color=genus,
           shape=water_origin))+
  geom_point(aes(alpha=0.5,
                 size=1))+
  xlab("Providencia sp. Isolate ID")+
  ylab("Prophage Length (bp)")+
  guides(alpha = "none",
         size="none",
         shape="none")

```


Comparing ENTEROBACTER sp.:
```{r}
entero_fam <- prophage_plot2%>%
  filter(family =="Enterobacteriaceae")%>%
  filter(is_virus =="yes")


ggplot(entero_fam,
       aes(x=isolateID,
           y=bp_length,
           color=genus,
           shape=water_origin))+
  geom_point(aes(alpha=0.5,
                 size=1))+
  xlab("Enterobacteriaceae Members Isolate ID")+
  ylab("Prophage Length (bp)")+
  guides(alpha = "none",
         size="none")

```

